---
title: "Class Exercise 2"
author: "Bernhard, Josia, Nico"
date: "`r Sys.Date()`"
output: html_document
---
# Number of Articles in the Guardian on Covid-19 and the Russian Invasion of Ukraine

This is the Class exercise 2 of Bernhard, Nico and Josia for the Data Mining in R class of spring 2024.

## Tasks
1) Register to the Guardian API and get the API key. 
2) Create a corpus of articles of choice following your instinct and interests.
3) Once you created your corpus, make a basic application of choice. 

## Hypothesis
The decline of articles on the topic of covid-19 goes along with a rise in articles on the topic of the russian invasion on Ukraine and with a decline of deaths of covid-19 in the UK.

## Method
We will use the Guardian API to extract the number of articles published on the topics of COVID-19 and the Russian invasion of Ukraine. We will then compare the number of articles published on these topics over time. As an extra we will add the weekly deaths in the UK due to COVID-19 to the analysis, which we will extract from the UK National Health Security Agency (NHS) API.

## Code

### Load libraries, clear workspace

```{r setup, eval = FALSE}
rm(list = ls())

library(httr)
library(jsonlite)
library(lubridate)
library(tidyverse)
```

### The Guardian data extraction

```{r eval = FALSE}
# load api key and base_url, set to search to find content
api_key <- rstudioapi::askForPassword("Enter Guardian API Key")
base_url <- "https://content.guardianapis.com/search"

# define the start and end date, both are chosen as a monday
start_date <- as.Date("2021-01-01")
end_date <- as.Date("2024-04-05")


# test whether we can get every start and end date of each week for the for loop later
week_dates <- seq(from = start_date, to = end_date, by = "week")

for(week_start in week_dates) {
  print(as.Date(week_start))
  
  end_date = week_start +7
  print(as.Date(end_date))
  cat("\n")
}


# init df to store covid results
results_covid <- data.frame(WeekStart = as.Date(character()), TotalArticles = integer())

# Loop over each week
for (week_start in week_dates) {


  week_end = week_start + 7
  
  # Set up API parameters for the current week
  params <- list(
    'api-key' = api_key,
    'q' = "COVID-19 OR coronavirus",
    'from-date' = as.character(as.Date(week_start)),
    'to-date' = as.character(as.Date(week_end))
  )
  
  # Perform the API call
  response <- GET(base_url, query = params)
  if (status_code(response) == 200) {
    content <- content(response, "text")
    json <- fromJSON(content)
    total_articles <- json$response$total
    
    # Append the results to the dataframe
    results_covid <- rbind(results_covid, data.frame(WeekStart = as.Date(week_end), TotalArticles = total_articles))
    
    # Debug output
    print(paste("Week starting:", as.Date(week_start), "- Total Articles:", total_articles))
  } else {
    print(paste("Failed to retrieve data for week starting:", as.Date(week_start)))
    
    
  }
  
  Sys.sleep(abs(rnorm(n = 1, mean = 1.5, sd = 0.4)))
}  


save(results_covid, file = "data/article_counts_covid.RData")

# init df to store russian/ukraine results
results_russia <- data.frame(WeekStart = as.Date(character()), TotalArticles = integer())

# Loop over each week
for (week_start in week_dates) {
  
  
  week_end = week_start + 7
  
  # Set up API parameters for the current week
  params <- list(
    'api-key' = api_key,
    'q' = "Russia OR Ukraine",
    'from-date' = as.character(as.Date(week_start)),
    'to-date' = as.character(as.Date(week_end))
  )
  
  # Perform the API call
  response <- GET(base_url, query = params)
  if (status_code(response) == 200) {
    content <- content(response, "text")
    json <- fromJSON(content)
    total_articles <- json$response$total
    
    # Append the results to the dataframe
    results_russia <- rbind(results_russia, data.frame(WeekStart = as.Date(week_end), TotalArticles = total_articles))
    
    # Debug output
    print(paste("Week starting:", as.Date(week_end), "- Total Articles:", total_articles))
  } else {
    print(paste("Failed to retrieve data for week starting:", as.Date(week_end)))
    
    
  }
  
  Sys.sleep(abs(rnorm(n = 1, mean = 1.5, sd = 0.4)))
}  

save(results_russia, file = "data/article_counts_russia.RData")

```

## NHS data extraction

```{r, eval = FALSE}

url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics"

response <- GET(url)
content <- content(response, "text")
json_overview <- fromJSON(content)




# try: COVID-19_cases_rateRollingMean -------------------------------------

url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_cases_rateRollingMean"

response <- GET(url)
content <- content(response, "text")
json <- fromJSON(content)

# this gives us just one datapoint, lets add filters at the end to get more


# try to get the weekly deaths for 365 pages ------------------------------

url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_deaths_ONSRegByWeek?page_size=365"

response <- GET(url)
content <- content(response, "text")
json <- fromJSON(content)

# this gives us the covid-19 deaths, but for some reason the data starts in 2021

# lets try the positivity rate, it is weekly aswell -----------------------

url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_cases_lineagePercentByWeek?page_size=365"
response <- GET(url)
content <- content(response, "text")
json <- fromJSON(content)

# this variables is actually the percentage of the prevalent variants of covid-19, not the positivity rate of tests


# lets go back to the deaths and try to get the ones from 2020 ------------


url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_deaths_ONSRegByWeek?year=2020"

response <- GET(url)
content <- content(response, "text")
json <- fromJSON(content)



# to add more filters, we use the params list -----------------------------

base_url_deaths <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_deaths_ONSRegByWeek"

q_params <- list(
  page_size = 365,
  year = 2020
)

response <- GET(base_url_deaths, query = q_params)
content <- content(response, "text")
json <- fromJSON(content)

# if i include the year it returns an empty list, maybe they dont have it, or im doing some mistake


# get the weekly death starting in 2021 -----------------------------------

base_url_deaths <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/respiratory/topics/COVID-19/geography_types/Nation/geographies/England/metrics/COVID-19_deaths_ONSRegByWeek"

q_params <- list(
  page_size = 365
)

response <- GET(base_url_deaths, query = q_params)
content <- content(response, "text")
json <- fromJSON(content)

# clean the data, extract the counts and dates ----------------------------

class(json$results)

UK_weekly_death <- json$results

names(UK_weekly_death)
View(UK_weekly_death)

UK_weekly_death <- UK_weekly_death %>% 
  select(date, metric_value) %>% 
  rename(COVID_deaths_UK = metric_value) %>% 
  mutate(date = as.Date(date)) %>% 
  rename(WeekStart = date)

sapply(UK_weekly_death, class)


save(UK_weekly_death, file = "data/UK_weekly_covid_death.RData")
```

## Data Analysis 

```{r, eval = FALSE}
load(file = "data/article_counts_covid.RData")
load(file = "data/article_counts_russia.RData")
load(file = "data/UK_weekly_covid_death.RData")

results_covid <- results_covid %>% 
  rename(Articles_covid19 = TotalArticles)

results_russia <- results_russia %>% 
  rename(Articles_ukraine_russia = TotalArticles)

df_combined <- left_join(results_covid, results_russia, by = "WeekStart")
df_comb_covid <- left_join(results_covid, UK_weekly_death, by = "WeekStart")


df_long <- df_combined %>%
  pivot_longer(cols = c(Articles_ukraine_russia, Articles_covid19),
               names_to = "Variable",
               values_to = "Count")


df_long_covid <- df_comb_covid %>%
  pivot_longer(cols = c(Articles_covid19, COVID_deaths_UK),
               names_to = "Variable",
               values_to = "Count")


plot_articles <- ggplot(data = df_long, aes(x = WeekStart, y = Count, color = Variable)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(title = "The Guardian Weekly Articles Published \non Ukraine/Russia and COVID-19", x = "Monthly Data", y = "Count", color = "Variable") +
  scale_color_manual(values = c("Articles_ukraine_russia" = "blue", "Articles_covid19" = "red")) +
  theme_minimal()


plot_covid <- ggplot(data = df_long_covid, aes(x = WeekStart, y = Count, color = Variable)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  labs(title = "The Guardian Weekly Articles Published \non COVID-19 and weekly UK COVID-19 deaths", x = "Monthly Data", y = "Count", color = "Variable") +
  scale_color_manual(values = c("Articles_ukraine_russia" = "blue", "COVID_deaths_UK" = "red")) +
  theme_minimal()

```

## Plots

```{r, eval = FALSE}

print(plot_articles)
print(plot_covid)

```

## Results and Discussion
 
The data shows that the number of articles published on the topics of COVID-19 and the Russian invasion of Ukraine have been fluctuating over time. The number of articles published on COVID-19 has been decreasing over time, while the number of articles published on the Russian invasion of Ukraine has been increasing. The number of weekly deaths in the UK due to COVID-19 has also been decreasing over time. 

Interestingly, there is an overlap between the increase in the number of articles published on the Russian invasion of Ukraine and the decrease in the number of articles published on COVID-19. This suggests that the decrease in the number of articles published on COVID-19 may be related to the increase in the number of articles published on the Russian invasion of Ukraine.

Furthermore, there is an overlap between the decrease in the number of articles published on COVID-19 and the decrease in the number of weekly deaths in the UK due to COVID-19. This suggests that the decrease in the number of articles published on COVID-19 may be related to the decrease in the number of weekly deaths in the UK due to COVID-19.

## Conclusion

Overall, the data suggests that the number of articles published on COVID-19 and the Russian invasion of Ukraine may be related to each other and to the number of weekly deaths in the UK due to COVID-19. Further analysis is needed to determine the exact nature of these relationships.



